On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-01-20 MS.tex
Small editing in abstract Add comments in footnotes for theoritical expectation Small rewriting of results (add results for SLA RGR vs AGR in mass) Rewriting of "measurements under 'ideal' conditions Move paragraph about stage representativity among traits as a new subsection in results

% Todo:
% Ditch abbrvos for SLA, wood density, Hmax?
% Fix plots
%   -- major revision of code -- > lots of repetition, hard to follow
%   -- Figure 2: add * against traits used in meta-analysis
% - seach on XXX, footnote
% - check refs to appendices in main text

@@

\item We propose that traits should influence growth in a way that depends on the size of individual plants. We outline mechanisms and hypotheses based on new theoritical work, and test these predictions in tree species using a meta-analysis of 103 studies ($>$ 500 correlations) for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate). We also recorded data for a list of 14 others traits commonly used in the trait literature. To capture the effects of plant size, we tested for a shift in the direction of correlation between growth rates and traits across three ontogenic stages: seedling, sapling and adult.

@@

\item Results were consistent with predictions, even with some limitations arising from unequal numbers of observation across stages. Specific leaf area was correlated with relative growth rate (RGR) in seedlings but not in adult plants. Correlations of growth with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with RGR only in one plant stage category: seedlings and adults, respectively. 
\item Despite a sizeable literature on traits and growth, major knowledge gaps across size-classes still exist; for one-third of traits a majority of correlations with growth reported in the literature ($>$ 75\%) come from a single and unique size-classe. Clearly, there is scope for strengthening our understanding on the link between traits and growth through adding further empirical studies for traits and size-classes underrepresented in the literature.

@@

Based on the above model and this single assumption about growth, we outline five hypotheses, which arise by considering the influence of trait variation on elements of equations  \ref{eq:eq0}-\ref{eq:eq4} (see Table \ref{tab:trade-offs}) and how these may interact with plant size. These hypotheses reflect a direct mechanistic causation between traits and growth rate. It is possible that other correlations may arise indirectly. Nevertheless, for purposes of this paper we adopt working hypotheses arising only from direct mechanistic causation. With only XXX exception, the predicted effects of the trait are also similar when growth is expressed on relative of absolute terms, and as diameter, height or mass growth (Table \ref{tab:trade-offs} (TRUE XXXX?).
\footnote{The correlation with RGR$_{\rm mass}$ is negative but becomes positive with AGR$_{\rm mass}$. For SLA, WD, Hmax, and Aarea, the direction of the predicted relationship is not affected by the growth measurement used. However RGR and AGR are fundamentally different for a lot of ecologists. So we adopted a two steps approach, first we were conservative by focusing exclusively on RGR, then we compared results with those of a more inclusive dataset with both RGR and AGR.}

@@

We aimed to identify studies reporting a correlation across species, between functional traits and growth rate (e.g. relative or absolute growth, in diameter, height or mass, see Table \ref{tab:trade-offs}. Systematic searches were conducted across the two most comprehensive bibliographic platforms: Google scholar and Web of Science (Thomson Reuters). While Web of Science guarantees a relatively stable and reproducible search environment, with clearly defined lists of indexed journals, its searches are limited to meta-data (i.e. title, keywords, and abstracts, \citealt{Beckmann:2012hn}). This limitation does not apply to Google Scholar. We supplemented these records with results from our own bibliographic libraries and suggestions from other researchers.

@@

\subsubsection*{Measurements under ideal conditions}\label{ideal-vs-entire-dataset}

The entire dataset included studies differing in the growth measurement used (RGR or AGR), the number of species, the experimental conditions (stressed vs unstressed), and the concordance between stage used to measure trait and growth rate. The main results reported here were obtained with an ``ideal'' subset of data. This ``ideal'' subset, included only studies recording RGR across \underline{\textgreater} 10 species, under unstressed conditions, and where growth rates and traits were measured on plants from the same stage (expect for Hmax and seed mass, where a particular stage is by definition required to measure the trait). These were then compare against results obtained with the entire dataset (including all noted correlations).

@@

\subsection*{Main effects}

@@

We now report results from the entire dataset, to determine if additional data would alter or reinforce our conclusions, drawn sometimes with limited replicates. This entire dataset included correlations measured with both AGR and RGR (Fig. \ref{fig:fig4}), across at least two species (but measured using several individuals), and did not control for the source of trait or growth (i.e. trait could be measured on a different stage) or environmental conditions. The relaxed criteria considerably increased the number of replicates, from $n = 137$ to $n = 267$ correlations across 93 articles. However, the additional replicates were unequally distributed across trait-stage combinations. Some trait-stage combinations already with a lot of data got even more support (e.g. from 25 to 59 correlation coefficients reported for SLA at seedling stage), while others still had few replicates. On the 13 traits-stage combinations studied in total, six had $\le$ five correlation coefficients: (i.e. WD-seedling, Hmax-sapling, Hmax-seedling, Seed mass-sapling, Aarea-adult and Aarea-sapling.

@@

We had no hypotheses about a change in the correlation between trait and growth as AGR was used instead of RGR, except for seed mass. For seed mass, effects size differed significantly between RGR and AGR at seedling stage; as expected the correlation with RGR was positive and shifted to non significant with AGR (Fig \ref{fig:fig4}d; CI not overlapping). For other traits, effect size were generally consistent between RGR and AGR when both measurement were available. Some differences occured but they were not easily attributed to any single cause. For instance for WD, the correlation at juvenile stage shifted from negative with RGR to non significant with AGR (i.e CI included zero, Fig. \ref{fig:fig4}b). The correlation coefficients recorded for AGR came from \citet{Augspurger:1984ct} and unlike others they used wood density values from adult plants together with growth from juveniles.

In total, correlations had been reported for at least two different stages across 18 traits out of the 19 targeted (Fig. \ref{fig:fig1}). We had thought it likely that for some traits correlations would be reported only for one size class. But this limitation applied only for LAR and NARmass, where studies appeared confined to seedling stages. However, for a majority of traits the number of observations across stage was higthly unbalanced; for one-third of the 19 traits $>$ 75\% of correlations with growth reported come from a single and unique plant stage (Fig. \ref{fig:fig1}).

@@

@@

TODO: XXXX Something about limited amount of data
Despite a sizeable literature on traits and growth, there are major knowledge gaps on the link between trait and growth. The number of correlations reported for each 19 traits is generally unbalanced in favor of a particular stage, for historical or practical reasons. This means that the knowledge we have on this trait is often biased toward a particular size-classe. Additional data are specifically required in small plants (seedling and sapling) for WD and Hmax, and in large plants (adult and sapling) for Aarea and more broadly for all traits requiring a measure of whole plant biomass (e.g. LMR, LAR).

>>>>

% Todo:
% Ditch abbrvos for SLA, wood density, Hmax?
% Fix plots
%   -- major revision of code -- > lots of repetition, hard to follow
%   -- Figure 2: add * against traits used in meta-analysis: no
% - seach on XXX, footnote
% - check refs to appendices in main text

@@

\item We propose that traits should influence growth in a way that depends on the size of individual plants. We outline mechanisms and hypotheses based on new theoritical work, and test these predictions in tree species using a meta-analysis of 103 studies ($>$ 500 correlations) for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate). We also recorded data for 14 others traits commonly used in the trait literature. To capture the effects of plant size, we tested for a shift in the direction of correlation between growth rates and traits across three ontogenic stages: seedling, sapling and adult.

@@

\item Results were consistent with predictions, even with some limitations arising from unequal numbers of observation across stages. Specific leaf area was correlated with relative growth rate (RGR) in seedlings but not in adult plants. Correlations of growth with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with RGR only in one plant stage category: seedlings and adults, respectively. 
\item Despite a sizeable literature on traits and growth, major knowledge gaps across size-classes still exist; for one-third of traits the majority of correlations with growth ($>$ 75\%) come from a single and unique size-classe. Clearly, there is scope for strengthening our understanding on the link between traits and growth through adding further empirical studies for traits and size-classes underrepresented in the literature.

@@

Based on the above model and this single assumption about growth, we outline five hypotheses, which arise by considering the influence of trait variation on elements of equations  \ref{eq:eq0}-\ref{eq:eq4} (see Table \ref{tab:trade-offs}) and how these may interact with plant size. These hypotheses reflect a direct mechanistic causation between traits and growth rate. It is possible that other correlations may arise indirectly. Nevertheless, for purposes of this paper we adopt working hypotheses arising only from direct mechanistic causation. With only XXX exception, the predicted effects of the trait are also similar when growth is expressed on relative of absolute terms, and as diameter, height or mass growth (Table \ref{tab:trade-offs} (TRUE XXXX?).

\footnote{The correlation with RGR$_{\rm mass}$ is negative but becomes positive with AGR$_{\rm mass}$. For SLA, WD, Hmax, and Aarea, the direction of the predicted relationship is not affected by the growth measurement used. However RGR and AGR are fundamentally different for a lot of ecologists. So we adopted a two steps approach, first we were conservative by focusing exclusively on RGR, then we compared results with those of a more inclusive dataset with both RGR and AGR.}
\footnote{We have different expectation for SLA in AGRmass for seedling: but none of our data are in AGRmass, they are mainly in RGRmass (Fig. \ref{fig:figA8}). Why a change with WD when it is measured in mass? All WD seedling are measured in RGRmass, and there are some mass data for both other stages too. }

@@

We aimed to identify studies reporting a correlation across species, between traits and growth rate (e.g. relative or absolute growth, in diameter, height or mass, see Table \ref{tab:trade-offs}). Systematic searches were conducted across the two most comprehensive bibliographic platforms: Google scholar and Web of Science (Thomson Reuters). While Web of Science guarantees a relatively stable and reproducible search environment, with clearly defined lists of indexed journals, its searches are limited to meta-data (i.e. title, keywords, and abstracts, \citealt{Beckmann:2012hn}). This limitation does not apply to Google Scholar. We supplemented these records with results from our own bibliographic libraries and suggestions from other researchers.

@@

\subsubsection*{Measurements under 'ideal' conditions}\label{ideal-vs-entire-dataset}

The entire dataset included studies differing in the growth measurement used (RGR or AGR), the number of species, the experimental conditions (stressed vs unstressed), and the concordance between stages used to measure trait and growth rate. However RGR and AGR are fundamentally different for a lot of ecologists. And the number of species used to perform the correlation, the experimental conditions, and the concordance between stages used to measure trait and growth are clear criteria for assessing the quality of the data for addressing our hypotheses. So we adopted a two steps approach, first we were conservative by focusing exclusively on RGR, across \underline{\textgreater} 10 species, under unstressed conditions, and where the source of the trait and growth is the same (expect for Hmax and seed mass, where a particular stage is by definition required to measure the trait). This dataset is called the "ideal" dataset. Then we compared results with those of a more inclusive dataset - the entire dataset- with both RGR and AGR, and all others criteria relaxed.

@@

\subsection*{Main effects under 'ideal' conditions}

@@

We now report results from the entire dataset to determine if additional data would alter or reinforce our conclusions, drawn sometimes with limited replicates. This entire dataset included correlations measured with both AGR and RGR (Fig. \ref{fig:fig4}), across at least two species (but measured using several individuals), and did not control for the source of trait or growth (i.e. trait could be measured on a different stage) or environmental conditions. The relaxed criteria considerably increased the number of replicates, from $n = 137$ to $n = 267$ correlations across 93 articles. However, the additional replicates were unequally distributed across trait-stage combinations. Some trait-stage combinations already with a lot of data got even more support (e.g. from 25 to 59 correlation coefficients reported for SLA at seedling stage), while others still had few replicates. On the 13 traits-stage combinations studied in total, six had $\le$ five correlation coefficients: WD-seedling, Hmax-sapling, Hmax-seedling, Seed mass-sapling, Aarea-adult and Aarea-sapling.

@@

We had hypotheses for seed mass and SLA, about a change in the correlation between trait and growth as AGR was used instead of RGR, or as mass was used instead of dimater and height. For seed mass, effects size differed significantly between RGR and AGR at seedling stage; as expected the correlation with RGR was positive and shifted to non significant with AGR (Fig \ref{fig:fig4}d; CI not overlapping). For SLA, the correlation was expected to shift from positive to non significant if AGRmass was used instead of RGRmass for seedling stage; here none of the data have been measured with AGRmass at seedling stage (Fig. \ref{fig:figA8}). 
% Why a change with WD when it is measured in mass? All WD seedling are measured in RGRmass, and there are some mass data for both other stages too.  
For other traits, effect size were generally consistent between RGR and AGR when both measurement were available. Some differences occured but they were not easily attributed to any single cause. For instance for WD, the correlation at juvenile stage shifted from negative with RGR to non significant with AGR (i.e CI included zero, Fig. \ref{fig:fig4}b). The correlation coefficients recorded for AGR came from \citet{Augspurger:1984ct} and unlike others they used wood density values from adult plants together with growth from juveniles.

@@

\subsection*{Size-classes representativity among 19 traits recorded}
In total, correlations had been reported for at least two different stages across 18 traits out of the 19 targeted (Fig. \ref{fig:fig1}). We had thought it likely that for some traits correlations would be reported only for one size class. But this limitation applied only for LAR and NARmass, where studies appeared confined to seedling stages. However, for a majority of traits the number of observations across stage was higthly unbalanced; for one-third of the 19 traits $>$ 75\% of correlations with growth reported come from a single and unique plant stage (Fig. \ref{fig:fig1}).

@@

TODO: XXXX Something about limited amount of data (add the following arguments:)
Despite a sizeable literature on traits and growth, there are major knowledge gaps on the link between trait and growth. The number of correlations reported for each 19 traits is generally unbalanced in favor of a particular stage, either for historical or practical reasons. This means that the knowledge we have on each trait is often biased toward a particular size-classe. Additional data are specifically required in small plants (seedling and sapling) for WD and Hmax, and in large plants (adult and sapling) for Aarea and more broadly for all traits requiring a measure of whole plant biomass (e.g. LMR, LAR).
